### The R markdown code for running GO enrichment on the pangenome
### Input: list of core genes, cloud genes and all genes, as well as the best 
### annotations file
### Output: GO enrichment dot plots
---
title: "Cluster1"
output: html_document
date: "2026-02-18"
---
install clusterprofiler and load packages
```{r}
library(clusterProfiler)
library(enrichplot)
library(tidyverse)  
packageVersion("clusterProfiler")
```



#Read in the data

```{r}
data <-read.delim("Best_annotations_16") #this is in the correct format col 1 gene, col2 GO ID
core_list <-read.table("Core_genes_3D7_cluster_sixteen.txt", header=FALSE)$V1 #make a vector of all the genes
cloud_list <-read.table("Cloud_genes_3D7_cluster_sixteen.txt", header=FALSE)$V1
all_list <- read.table("allgenes_3D7_cluster_sixteen.txt", header=FALSE)$V1

```

#Filter for the first two columns(TERM2GENE)
#Filter  for first and third (TERM2NAME)


```{r}
cc_t2g <- data%>%filter(Type=="CC")%>%select(GO_ID,Gene_ID)
bp_t2g <-data %>%filter(Type=="BP")%>%select(GO_ID,Gene_ID)
mf_t2g <-data %>%filter(Type=="MF")%>%select(GO_ID, Gene_ID)
cc_t2n <- data%>%filter(Type=="CC")%>%select(GO_ID,GO_term)
bp_t2n <-data %>%filter(Type=="BP")%>%select(GO_ID,GO_term)
mf_t2n <-data %>%filter(Type=="MF")%>%select(GO_ID, GO_term)
```


#perform analysis for cc
```{r}
comparison_cc<-compareCluster(geneClusters=list(core=core_list,cloud=cloud_list),
          fun="enricher",
          TERM2GENE=cc_t2g,
          TERM2NAME=cc_t2n,
          pvalueCutoff=0.05,
          pAdjustMethod="BH",
          universe=all_list,
          minGSSize=5, #more lenient
          maxGSSize=500,#default
          qvalueCutoff=0.2)#default

          
```

#visualise and save
```{r}
p<-dotplot(comparison_cc, showCategory=15)
ggsave("GO_CC_dotplot_sixteen.png", 
       plot = p,
       width = 15, 
       height = 30, 
       units = "cm")
```


#perform analysis for bp
```{r}
comparison_bp<-compareCluster(geneClusters=list(core=core_list,cloud=cloud_list),
          fun="enricher",
          TERM2GENE=bp_t2g,
          TERM2NAME=bp_t2n,
          pvalueCutoff=0.05,
          pAdjustMethod="BH",
          universe=all_list,
          minGSSize=5, #more lenient
          maxGSSize=500,#default
          qvalueCutoff=0.2)#default

          
```

#visualise
```{r}
p2<-dotplot(comparison_bp, showCategory=15)
ggsave("GO_BP_dotplot_sixteen.png", 
       plot = p2,
       width = 15, 
       height = 30,  
       units = "cm")


```


#perform analysis for mpf
```{r}
comparison_mf<-compareCluster(geneClusters=list(core=core_list,cloud=cloud_list),
          fun="enricher",
          TERM2GENE=mf_t2g,
          TERM2NAME=mf_t2n,
          pvalueCutoff=0.05,
          pAdjustMethod="BH",
          universe=all_list,
          minGSSize=5, #more lenient
          maxGSSize=500,#default
          qvalueCutoff=0.2)#default

          
```

#visualise
```{r}
p3<-dotplot(comparison_mf, showCategory=15)
ggsave("GO_MF_dotplot_sixteen.png", 
       plot = p3,
       width = 15, 
       height = 30, 
       units = "cm")
```





